---
title: 
author: "Name: Ruixuan Li"
- |
    | Student number: 24221917
date: "`r format(Sys.time(), '%X, %d %B, %Y')`"
output: html_document
---
# Originality declaration  

I, [Ruixuan Li], confirm that the work presented in this assessment is my own. Where information has been derived from other sources, I confirm that this has been indicated in the work.

date: `r format(Sys.time(), '%d %B, %Y')`

Introduction


# Start your response here
```{r}
# load libraries
library(readr)
library(ggplot2)
library(sf)
library(tidyverse)
library(spdep)
library(spatstat)
library(sp)
library(tmap)
library(spdep)
library(janitor)
```
research question:hat are the factors that might lead to variation in Graffiti occurrence across the city? 



```{r}
households_families <- read_csv("/Users/liruixuan/Desktop/UCL/GIS/casa0005-practice-exam-2024-25-lrxzyx-main/\ Households\ and\ Families\ .csv")

poverty_data <- read_csv("/Users/liruixuan/Desktop/UCL/GIS/casa0005-practice-exam-2024-25-lrxzyx-main/povety.csv")

graffiti_points <- read_csv("/Users/liruixuan/Desktop/UCL/GIS/casa0005-practice-exam-2024-25-lrxzyx-main/Graffiti\ points.csv")

census_tract_data <- 
read_csv("/Users/liruixuan/Desktop/UCL/GIS/casa0005-practice-exam-2024-25-lrxzyx-main/Census\ tract\ spatial\ data.csv") 

education_data <- read_csv("/Users/liruixuan/Desktop/UCL/GIS/casa0005-practice-exam-2024-25-lrxzyx-main/Educational\ Attainment.csv")
```

```{r}
#join the factor data
factor_data <- left_join(households_families,poverty_data, education_data,by=c("GEO_ID"="GEO_ID"))
```

```{r}
colnames(factor_data)

```

```{r}
#join all the data
factor_census <- left_join(census_tract_data,factor_data,
            by = c("Geographic Area Name (NAME)" = "NAME.x"))

```

```{r}
colnames(factor_census)
```
```{r}
#rename the colname
factor_census <- factor_census %>% 
  rename("Total households"="S1101_C01_001E",
        "Average household size"="S1101_C01_002E",
          "Geographic Area Name"="NAME.y",
          "Population for whom poverty status is determined"="S1701_C01_001E")
```


```{r}
library(stringr)

# Extract latitude and longitude and create a new column
graffiti_points <- graffiti_points %>%
  mutate(
    longitude = str_extract(Point, "(?<=\\().*?(?=\\,)") %>% as.numeric(),
    latitude = str_extract(Point, "(?<=\\, ).*?(?=\\))") %>% as.numeric()
  )

#(?<=\\().*?(?=\\,) Extract the part between the left parenthesis (and comma) from the string (?<=...)(Lookbehind)  (?=...) (Lookahead)
```

```{r}
graffiti_points <- graffiti_points %>%
  filter(!is.na(longitude) & !is.na(latitude))
```


```{r}
#visualise graffiti_points
library(sf)

# 转换为 sf 对象
graffiti_sf <- st_as_sf(
  graffiti_points,
  coords = c("longitude", "latitude"),
  crs = 4326
)

# 查看转换结果
print(graffiti_sf)

```
```{r}
any(is.na(as.numeric(factor_census$Longitude)))
any(is.na(as.numeric(factor_census$Latitude)))

factor_census <- factor_census %>%
  filter(!is.na(Longitude) & !is.na(Latitude))

factor_census <- factor_census %>%
  mutate(
    Longitude = as.numeric(gsub("−", "-", Longitude)),  # 替换特殊负号并转换为数值
    Latitude = as.numeric(gsub("−", "-", Latitude))    # 如果 Latitude 也有特殊负号，可以一并处理
  )


```
```{r}
factor_census <- st_as_sf(factor_census, coords = c("Longitude", "Latitude"), crs = 4326)
```

```{r}
# 检查涂鸦点范围
st_bbox(graffiti_sf)

# 检查普查区域范围
st_bbox(factor_census)
```



```{r}
overlap <- st_within(graffiti_sf, factor_census)
if (!any(lengths(overlap) > 0)) {
  print("No graffiti points fall within census tracts.")
}
```


```{r}
# 计算每个普查区域的涂鸦数量
graffiti_count <- graffiti_within_tracts %>%
  group_by(CaseID) %>%
  summarise(graffiti_count = n())
```


```{r}
# 将涂鸦数量加入综合数据集
merged_data <- merged_data %>%
  left_join(graffiti_count, by = "tract_id") %>%
  mutate(graffiti_count = replace_na(graffiti_count, 0))

# 可视化涂鸦热点
ggplot(merged_data) +
  geom_sf(aes(fill = graffiti_count)) +
  scale_fill_viridis_c() +
  labs(title = "Graffiti Hotspots in San Francisco",
       fill = "Graffiti Count") +
  theme_minimal()
```

```{r}
# 简单回归分析
model <- lm(graffiti_count ~ poverty_rate + education_level, data = merged_data)
summary(model)

# 可视化关系
ggplot(merged_data, aes(x = poverty_rate, y = graffiti_count)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Graffiti Count vs Poverty Rate",
       x = "Poverty Rate (%)",
       y = "Graffiti Count") +
  theme_minimal()
```


```{r}
ggplot() +
  geom_sf(data = census_spatial, fill = "white", color = "black") +
  geom_sf(data = graffiti_sf, color = "red", size = 1) +
  labs(title = "Graffiti Incidents in San Francisco",
       subtitle = "Based on 2004 Data") +
  theme_minimal()
```

```{r}
ggplot(merged_data) +
  geom_sf(aes(fill = poverty_rate)) +
  geom_point(data = graffiti_sf, aes(geometry = geometry), color = "red", size = 1) +
  scale_fill_viridis_c() +
  labs(title = "Poverty Rate and Graffiti Incidents",
       fill = "Poverty Rate (%)") +
  theme_minimal()
```



## Initial project scope
6. 关键发现和结论
在最终报告中总结关键发现，例如：

涂鸦热点主要集中在某些区域。
贫困率和教育水平与涂鸦数量存在显著关联。
建议城市立法者采取主动措施，例如增加社区资源或加强高风险区域的监控。
