
# Originality declaration  

I, [**Ruixuan Li**], confirm that the work presented in this assessment is my own. Where information has been derived from other sources, I confirm that this has been indicated in the work.

# Start your response here
```{r load libraries}
library(readr)
library(ggplot2)
library(sf)
library(dplyr)
library(spdep)
library(spatstat)
library(sp)
library(tmap)
library(tmaptools)
library(janitor)
library(tidyverse)

```

```{r read data}
census_health <- read_csv("/Users/liruixuan/Desktop/UCL/GIS/examination/census2021-ts037/census2021-ts037-lsoa.csv",na=c(" "))
boundary <-st_read("/Users/liruixuan/Desktop/UCL/GIS/examination/Lower_layer/Lower_layer_Super_Output_Areas_December_2021_Boundaries_EW_BFC_V10.shp")
```

data
 LSOA (Lower Layer Super Output Area) for both census_health and boundary
boundary data is too large to upload :https://geoportal.statistics.gov.uk/datasets/2bbaef5230694f3abae4f9145a3a9800_0/explore (shapfile) 

```{r}

```

## Initial project scope