
# Originality declaration  

I, [**Ruixuan Li**], confirm that the work presented in this assessment is my own. Where information has been derived from other sources, I confirm that this has been indicated in the work.

# Start your response here

```{r load libraries}
library(readr)
library(ggplot2)
library(sf)
library(dplyr)
library(spdep)
library(spatstat)
library(sp)
library(tmap)
library(tmaptools)
library(janitor)
library(tidyverse)
```

```{r data reading}
census_health <- read_csv("/Users/liruixuan/Desktop/UCL/GIS/examination/examination/census2021-ts037-lsoa.csv",na=c(" "))

boundary <- st_read("/Users/liruixuan/Desktop/UCL/GIS/examination/examination/Lower_layer_Super_Output_Areas_December_2021_Boundaries_EW_BFC_V10/Lower_layer_Super_Output_Areas_December_2021_Boundaries_EW_BFC_V10.shp")

```
data:
LSOA (Lower Layer Super Output Area)
boundary data is too large to upload:
https://geoportal.statistics.gov.uk/datasets/2bbaef5230694f3abae4f9145a3a9800_0/explore  (shapefile) I removed all parentheses from the file name for easy reading of data

```{r}
#View the first few lines of data to understand the structure
head(census_health)
```

```{r Merge two data}
merged_data <- boundary %>%
  left_join(census_health, by = c("LSOA21CD" = "geography code"))
```

```{r filter data}
#Filter out Manchester data
manchester_data <- merged_data %>%
  filter(grepl("Manchester", LSOA21NM, ignore.case = TRUE))
```

```{r tmap}
#Create a map
tm_shape(manchester_data) +
  tm_polygons("General health: Total: All usual residents", 
              title = "Health Status in Manchester") +
  tm_layout(main.title = "Health in Manchester")
```

```{r}
window <- as.owin(manchester_data) 
plot(window)
```
```{r create a sp objec}

```

## Initial project scope